---
title: "Real value of different portfolios to reach $1m in 2020"
output: pdf_document
---

```{r setup, include=FALSE}
rm(list=ls())
library(dplyr)
library(ggplot2)
library(tidyr)
library(lubridate)
library(reshape2)

return = read.csv("datasets/Inflation_Adjusted_Return.csv")
rob = read.csv("datasets/Robert_Shiller_CPI_adjusted_Return.csv")
limit = read.csv("datasets/IRA-Contribution-Limit.csv")
return = merge(return, rob, by=c("Year","Month"))
colnames(return)[18:19] = c("RS_adjusted_Price","RS_adjusted_GR")
# View(return)
```

### Inflation measures: CPI, CPIC, PPI, PCE
### Return types: GR(gross return), PR(price return)

```{r, echo=FALSE}
## Add&adjust Time&portfolio column
measures = c("CPI", "CPIC", "PPI", "PCE")
return_types = c("GR", "PR") ##GR: gross return, PR: price return
for (measure in measures){
  for (return_type in return_types){
    return[nrow(return),ncol(return)+1] = 1000000 ## To reach 1m after 2019 Dec
    names(return)[ncol(return)] = paste0(measure,"_",return_type,"_","portfolio")
  }
}
return = return[seq(dim(return)[1],1),]
for (i in 20:27){
  for (j in 2:nrow(return)){
    if (i==20 | i==21){ return[j,i]=return[j-1,i]/(1+return[j,i-14]) }
    if (i==22 | i==23){ return[j,i]=return[j-1,i]/(1+return[j,i-13]) }
    if (i==24 | i==25){ return[j,i]=return[j-1,i]/(1+return[j,i-12]) }
    if (i==26 | i==27){ return[j,i]=return[j-1,i]/(1+return[j,i-11]) }
  }
}

## Robert Shiller Adjusted returns
return[1,ncol(return)+1] = 1000000 ## To reach 1m after 2019 Dec
names(return)[ncol(return)] = "RS_GR_portfolio"
for (j in 2:nrow(return)){ return[j,28]=return[j-1,28]/(1+return[j,19]) }

return = return %>% mutate(Time=paste0(Year,'/',sprintf("%02d", return$Month),'/',01))
return$Time = as.Date(return$Time, format = "%Y/%m/%d")
return = return[seq(dim(return)[1],1),]
return = na.omit(return)
```


```{r, fig.height=3.5, fig.width=8, echo=FALSE, warning=FALSE}
# par(mfrow = c(2, 2))
# reshape data from wide to long format
# GR normal scale
long = melt(return, measure.vars=c('CPI_GR_portfolio','CPIC_GR_portfolio','PPI_GR_portfolio','RS_GR_portfolio')) %>% select(Time, variable, value)
g = ggplot(long, aes(Time, value, shape = variable, color = factor(variable))) + 
  geom_line() +
  scale_x_date(date_breaks = "10 years")+
  labs(title="Starting GR portfolio real value to reach $1m in 2020", color = "Portfolio")+
  ylab("Portfolio real value in US dollars")+
  guides(color = guide_legend(override.aes = list(size=3)))+
  theme_minimal()
print(g)

# PR normal scale
long = melt(return, measure.vars=c('CPI_PR_portfolio','CPIC_PR_portfolio','PPI_PR_portfolio')) %>% select(Time, variable, value)
g = ggplot(long, aes(Time, value, shape = variable, color = factor(variable))) + 
  geom_line() +
  scale_x_date(date_breaks = "10 years")+
  labs(title="Starting PR portfolio real value to reach $1m in 2020", color = "Portfolio")+
  ylab("Portfolio real value in US dollars")+
  guides(color = guide_legend(override.aes = list(size=3)))+
  theme_minimal()
print(g)

## Log-scaled portfolios
for (col in 20:28){return[col] = log(return[col])}
# reshape data from wide to long format
# GR logged scale
long = melt(return, measure.vars=c('CPI_GR_portfolio','CPIC_GR_portfolio','PPI_GR_portfolio','RS_GR_portfolio')) %>% select(Time, variable, value)
g = ggplot(long, aes(Time, value, shape = variable, color = factor(variable))) + 
  geom_line() +
  scale_x_date(date_breaks = "10 years")+
  labs(title="Log starting GR portfolio real value to reach $1m in 2020", color = "Portfolio")+
  ylab("Log portfolio real value in US dollars")+
  guides(color = guide_legend(override.aes = list(size=3)))+
  theme_minimal()
print(g)

# PR logged scale
long = melt(return, measure.vars=c('CPI_PR_portfolio','CPIC_PR_portfolio','PPI_PR_portfolio')) %>% select(Time, variable, value)
g = ggplot(long, aes(Time, value, shape = variable, color = factor(variable))) + 
  geom_line() +
  scale_x_date(date_breaks = "10 years")+
  labs(title="Log starting portfolio real value to reach $1m in 2020", color = "Portfolio")+
  ylab("Log portfolio real value in US dollars")+
  guides(color = guide_legend(override.aes = list(size=3)))+
  theme_minimal()
print(g)
```


```{r,echo=FALSE}
## Time taken from $0 to $1m
for (measure in measures){
  for (return_type in return_types){
    return[1,ncol(return)+1] = 0 ## To reach 1m after 2019 Dec
    names(return)[ncol(return)] = paste0(measure,"_",return_type,"_","t")
  }
}

## RS
return[1,ncol(return)+1] = 0 ## To reach 1m after 2019 Dec
names(return)[ncol(return)] = "RS_GR_t"

```

## Contribute $500 each month
```{r, eval=FALSE, include=FALSE}
ptm = proc.time()
for (col in 29:37){
  for (i in 1:nrow(return)){
    ct = 0
    portfolio_value = 0
    temp = i
    fulfiled = T
    while (portfolio_value<1000000){
      if (temp==1){
        portfolio_value = portfolio_value+500
        temp = temp+1
      } else if (temp>nrow(return)){
        fulfiled = F
        break
      } else {
        if (col==29|col==30){portfolio_value = (portfolio_value+500)*(1+return[temp-1,col-23])}
        else if (col==31|col==32){portfolio_value = (portfolio_value+500)*(1+return[temp-1,col-22])}
        else if (col==33|col==34){portfolio_value = (portfolio_value+500)*(1+return[temp-1,col-21])}
        else if (col==35|col==36){portfolio_value = (portfolio_value+500)*(1+return[temp-1,col-20])}
        else if (col==37){portfolio_value = (portfolio_value+500)*(1+return[temp-1,col-18])}
        temp = temp+1
        # print(paste('col:',col,'ct:',ct,'portfolio:', portfolio_value))
      }
      ct = ct+1
    }
    if (fulfiled){return[i,col] = ct} else {return[i,col] = NA}
    if (i%%120==0){print(paste("Portfolio:",col,"Time:",return$Time[i],"t:",ct, "Time elapsed:", as.numeric(proc.time()-ptm)[3]))}
  }
}

# save(return, file="Time to 1m.RData")

```


\newpage
## Distribution of time taken from $0 to $1m for different portfolio x deflation method combinations
Red vertical lines in histograms represents the mean time for a portfolio to reach $1m

```{r, echo=FALSE, warning=FALSE, fig.height=3.5, fig.width=8}
load("Time to 1m.RData")

for (col in c(29,30,31,32,33,34,37)){
  ## Convert t to year scale
  return[col]=return[col]/12
  if (col==29) {x1=-0.9;x2=-0.9;y1=60;y2=y1;pf='CPI_GR'}
  else if (col==30) {x1=-1.1;x2=-1.4;y1=26;y2=21;pf='CPI_PR'}
  else if (col==31) {x1=-0.8;x2=-0.8;y1=55;y2=y1;pf='CPIC_GR'}
  else if (col==32) {x1=-1;x2=2;y1=24;y2=y1;pf='CPIC_PR'}
  else if (col==33) {x1=-1.2;x2=1.2;y1=50;y2=y1;pf='PPI_GR'}
  else if (col==34) {x1=-1;x2=1;y1=28;y2=y1;pf='PPI_PR'}
  else if (col==37) {x1=-1;x2=1;y1=25;y2=y1;pf='RS_GR'}
  p=ggplot(return, aes_string(names(return)[col]))+
      geom_bar(fill="steelblue",color="white", binwidth = 1)+
      labs(title=paste('Distribution of time to reach $1m (in years) for',pf))+
      xlab("Time to reach $1m in years")+
      ylab('Time to reach $1m from different\nstart time (in years)')+
      geom_vline(aes(xintercept = mean(return[,col],na.rm = TRUE)),col='blue',size=2)+
      annotate("text", x=mean(return[,col],na.rm = TRUE)+x1, y=y1, label= paste("Mean:\n",round(mean(return[,col],na.rm = TRUE),1),"years"), size=3.5, col='blue')+
      geom_vline(aes(xintercept = median(return[,col],na.rm = TRUE)),col='red',size=2)+
      annotate("text", x=median(return[,col],na.rm = TRUE)+x2, y=y2, label= paste("Median:\n",round(median(return[,col],na.rm = TRUE),1),"years"), size=3.5, col='red')+
      theme_minimal()
  
  print(p)
}

```

## PCE_PR portfolio never reaches $1m from any starting time since 1959 Feb.


```{r, echo=FALSE, warning=FALSE, fig.height=3.5, fig.width=8}
# GR
return.temp = return %>% 
  gather(Portfolio, t, c('CPI_GR_t','CPIC_GR_t','PPI_GR_t','RS_GR_t')) %>%
  select(Portfolio, t)
p = ggplot(return.temp, aes(x=Portfolio, y=t), color=as.factor(Portfolio)) + 
    geom_boxplot(fill="white",color="steelblue")+
    xlab("Gross Return(GR) Portfolios")+
    ylab("Time to Reach $1m in Years")+
    theme_minimal()+
    labs(title="Boxplot of Time to Reach $1m for Different Gross Return(GR) Portfolios")+ 
    annotate("text", x=4, y=median(return.temp$t,na.rm = TRUE)-3, label= paste("Median:",round(median(return.temp$t,na.rm = TRUE),1),"years"),color="blue")+
    geom_hline(aes(yintercept = median(return.temp$t,na.rm = TRUE)),col='blue',lty=2,size=1)+
    stat_summary(fun.y=median, colour="red", geom="text", show_guide = FALSE, vjust=1.5, aes(label=round(..y.., digits=1)))+
    stat_summary(fun.y=median, colour="darkred", geom="point", size=3,show_guide = FALSE)+
    annotate("text", x=4, y=mean(return.temp$t,na.rm = TRUE)+3, label= paste("Mean:",round(mean(return.temp$t,na.rm = TRUE),1),"years"),color="red")+
    geom_hline(aes(yintercept = mean(return.temp$t,na.rm = TRUE)),col='red',lty=2,size=1)+
    scale_x_discrete(labels=c('CPI_GR','CPIC_GR','PPI_GR','RS_GR'))
print(p)

#PR
return.temp = return %>% 
  gather(Portfolio, t, c('CPI_PR_t','CPIC_PR_t','PPI_PR_t')) %>%
  select(Portfolio, t)
p = ggplot(return.temp, aes(x=Portfolio, y=t), color=as.factor(Portfolio)) + 
    geom_boxplot(fill="white",color="steelblue")+
    xlab("Price Return(PR) Portfolios")+
    ylab("Time to Reach $1m in Years")+
    theme_minimal()+
    labs(title="Boxplot of Time to Reach $1m for Different Price Return(PR) Portfolios")+ 
    annotate("text", x=3, y=median(return.temp$t,na.rm = TRUE)+2, label= paste("Median:",round(median(return.temp$t,na.rm = TRUE),1),"years"),color="blue")+
    geom_hline(aes(yintercept = median(return.temp$t,na.rm = TRUE)),col='blue',lty=2,size=1)+
    stat_summary(fun.y=median, colour="red", geom="text", show_guide = FALSE, vjust=1.5, aes(label=round(..y.., digits=1)))+
    stat_summary(fun.y=median, colour="darkred", geom="point",size=3,show_guide = FALSE)+
    annotate("text", x=3, y=mean(return.temp$t,na.rm = TRUE)-2, label= paste("Mean:",round(mean(return.temp$t,na.rm = TRUE),1),"years"),color="red")+
    geom_hline(aes(yintercept = mean(return.temp$t,na.rm = TRUE)),col='red',lty=2,size=1)+
    scale_x_discrete(labels=c('CPI_PR','CPIC_PR','PPI_PR'))
print(p)
```




\newpage

```{r,echo=FALSE, fig.height=3.5, fig.width=8, echo=FALSE, warning=FALSE}
# GR
unfulfiled_t = data.frame("Portfolio"='a',
                          "Unfulfiled.start.time"=as.Date("2019-12-01"),
                          "Unfulfiled.end.time"=as.Date("2019-12-01"),stringsAsFactors=FALSE)
for (col in c(29,31,33,37)){
  unfulfiled_t=rbind(list(as.factor(colnames(return)[col]),
                          as.Date(return$Time[which(is.na(return[col]))[1]]),
                          as.Date("2019-12-01")), unfulfiled_t)
}
unfulfiled_t=unfulfiled_t[-nrow(unfulfiled_t),]
ggplot() + 
  geom_linerange(data=unfulfiled_t, aes(x=Portfolio, ymin=Unfulfiled.start.time, ymax=Unfulfiled.end.time),size=1.5)+
  coord_flip()+
  xlab("Portfolio")+
  ylab("Start time")+
  theme_minimal()+
  labs(title="Impossible Start Time for GR Portfolios to Reach $1m in 2020")+ 
  annotate("text", y=min(unfulfiled_t$Unfulfiled.start.time)+900, x=3.7,
           label=min(unfulfiled_t$Unfulfiled.start.time),color="blue")+
  annotate("point", y=min(unfulfiled_t$Unfulfiled.start.time), x=4, colour = "lightblue", size=5)+
  annotate("text", y=max(unfulfiled_t$Unfulfiled.start.time)+900, x=1.7,
           label=max(unfulfiled_t$Unfulfiled.start.time),color="blue")+
  annotate("point", y=max(unfulfiled_t$Unfulfiled.start.time), x=2, colour = "lightblue", size=5)

#PR
unfulfiled_t = data.frame("Portfolio"='a',
                          "Unfulfiled.start.time"=as.Date("2019-12-01"),
                          "Unfulfiled.end.time"=as.Date("2019-12-01"),stringsAsFactors=FALSE)
for (col in c(30,32,34,36)){
  unfulfiled_t=rbind(list(as.factor(colnames(return)[col]),
                          as.Date(return$Time[which(is.na(return[col]))[1]]),
                          as.Date("2019-12-01")), unfulfiled_t)
}
unfulfiled_t=unfulfiled_t[-nrow(unfulfiled_t),]
ggplot() + 
  geom_linerange(data=unfulfiled_t, aes(x=Portfolio, ymin=Unfulfiled.start.time, ymax=Unfulfiled.end.time),size=1.5)+
  coord_flip()+
  xlab("Portfolio")+
  ylab("Start time")+
  theme_minimal()+
  labs(title="Impossible Start Time for PR Portfolios to Reach $1m in 2020")+ 
  annotate("text", y=min(unfulfiled_t$Unfulfiled.start.time)+1200, x=2.7,
           label=min(unfulfiled_t$Unfulfiled.start.time),color="blue")+
  annotate("point", y=min(unfulfiled_t$Unfulfiled.start.time), x=3, colour = "lightblue", size=5)+
  annotate("text", y=max(unfulfiled_t$Unfulfiled.start.time)+1200, x=3.7, 
           label=max(unfulfiled_t$Unfulfiled.start.time),color="blue")+
  annotate("point", y=max(unfulfiled_t$Unfulfiled.start.time), x=4, colour = "lightblue", size=5)

```






## Contribute based on historical contribution limit
```{r, eval=FALSE, include=FALSE}
load("Time to 1m.RData")
return = merge(return, limit, by="Year")

## Time taken from $0 to $1m
for (measure in measures){
  for (return_type in return_types){
    return[1,ncol(return)+1] = 0 ## To reach 1m after 2019 Dec
    names(return)[ncol(return)] = paste0(measure,"_",return_type,"_","t2")
  }
}
## RS
return[1,ncol(return)+1] = 0 ## To reach 1m after 2019 Dec
names(return)[ncol(return)] = "RS_GR_t2"

## inflation-adjusted contribution limit
for (measure in measures){
  return[1,ncol(return)+1] = 0 ## To reach 1m after 2019 Dec
  names(return)[ncol(return)] = paste0(measure,"_","limit")
}
return[1,ncol(return)+1] = 0
names(return)[ncol(return)] = paste0('RS',"_","limit")
return = return[seq(dim(return)[1],1),]
for (col in 48:52){
  for (i in 1:nrow(return)){
    if (col==48){rate=5}
    else if (col==49){rate=8}
    else if (col==50){rate=11}
    else if (col==51){rate=14}
    else if (col==52){rate=5}
    return[i,col]=return[i,38]*cumprod(1+return[rate])[i,1]
  }
}

```

```{r, eval=FALSE, include=FALSE}
return = return[seq(dim(return)[1],1),]
ptm = proc.time()
for (col in 39:47){
# for (col in 39){
  for (i in 1:nrow(return)){
  # for (i in 41:43){
    ct = 0
    portfolio_value = 0
    temp = i
    fulfiled = T
    while (portfolio_value<1000000){
      if (temp==1){
        if (col==39|col==40){portfolio_value = portfolio_value+return[i,48]/12}
        else if (col==41|col==42){portfolio_value = portfolio_value+return[i,49]/12}
        else if (col==43|col==44){portfolio_value = portfolio_value+return[i,50]/12}
        else if (col==45|col==46){portfolio_value = portfolio_value+return[i,51]/12}
        else if (col==47){portfolio_value = portfolio_value+return[i,52]/12}
        temp = temp+1
        # print(paste("Time:",return$Time[i],'ct:',ct,'portfolio:', portfolio_value))
      } else if (temp>nrow(return)){
        fulfiled = F
        break
      } else {
        if (col==39|col==40){portfolio_value = (portfolio_value+return[i,48]/12)*(1+return[temp-1,col-33])}
        else if (col==41|col==42){portfolio_value = (portfolio_value+return[i,49]/12)*(1+return[temp-1,col-32])}
        else if (col==43|col==44){portfolio_value = (portfolio_value+return[i,50]/12)*(1+return[temp-1,col-31])}
        else if (col==45|col==46){portfolio_value = (portfolio_value+return[i,51]/12)*(1+return[temp-1,col-30])}
        else if (col==47){portfolio_value = (portfolio_value+return[i,52]/12)*(1+return[temp-1,col-28])}
        temp = temp+1
        # print(paste('col:',col,'ct:',ct,'portfolio:', portfolio_value))
        # print(paste("Time:",return$Time[i],'ct:',ct,'portfolio:', round(portfolio_value,0), "Lim:", round(return[i,48]/12),0))
      }
      ct = ct+1
    }
    if (fulfiled){return[i,col] = ct} else {return[i,col] = NA}
    if (i%%12==0){print(paste("Portfolio:",col,"Time:",return$Time[i],"t:",ct, "Time elapsed:", as.numeric(proc.time()-ptm)[3]))}
  }
}

# name = c('CPI_GR','CPI_PR','CPIC_GR','CPIC_PR','PPI_GR','PPI_PR','PCE_GR','PCE_PR','RS_GR')
# names(return)[39:47] = paste0(name,"_","terminal.val")
save(return, file="Time to 1m-adj lim.RData")

```

\newpage
## Distribution of time taken from $0 to $1m for different portfolio x deflation method combinations
Red vertical lines in histograms represents the mean time for a portfolio to reach $1m

```{r, echo=FALSE, warning=FALSE, fig.height=3.5, fig.width=8}
load("Time to 1m-adj lim.RData")

for (col in c(39:44,47)){
  ## Convert t to year scale
  return[col]=return[col]/12
  space = '\n'
  if (col==39) {x1=-1;x2=-1;y1=25;y2=y1;pf='CPI_GR'}
  else if (col==40) {x1=-0.2;x2=-0.45;y1=21;y2=15;pf='CPI_PR'}
  else if (col==41) {x1=-1;x2=1;y1=30;y2=y1;pf='CPIC_GR'}
  else if (col==42) {x1=-0.25;x2=-0.17;y1=17;y2=y1;pf='CPIC_PR'}
  else if (col==43) {x1=-1.1;x2=1.1;y1=24.5;y2=y1;pf='PPI_GR'}
  else if (col==44) {x1=-0.35;x2=0.35;y1=24;y2=y1;pf='PPI_PR';space=''}
  else if (col==47) {x1=-0.3;x2=-0.45;y1=25;y2=19;pf='RS_GR'}
  p=ggplot(return, aes_string(names(return)[col]))+
      geom_bar(fill="steelblue",color="white", binwidth = 1)+
      labs(title=paste('Distribution of time to reach $1m in years for',pf))+
      xlab("Time to reach $1m in years")+
      ylab('Frequency')+
      geom_vline(aes(xintercept = mean(return[,col],na.rm = TRUE)),col='blue',size=2)+
      annotate("text", x=mean(return[,col],na.rm = TRUE)+x1, y=y1, label= paste("Mean:",space,round(mean(return[,col],na.rm = TRUE),1),"years"), size=3.5, col='blue')+
      geom_vline(aes(xintercept = median(return[,col],na.rm = TRUE)),col='red',size=2)+
      annotate("text", x=median(return[,col],na.rm = TRUE)+x2, y=y2, label= paste("Median:",space,round(median(return[,col],na.rm = TRUE),1),"years"), size=3.5, col='red')+
      theme_minimal()
  
  print(p)
}



```

## Inflation-adjusted Yearly Contribution Limit Changes
```{r, echo=FALSE, warning=FALSE, fig.height=3.5, fig.width=8}
plot(y=return$CPI_limit,x=return$Time,main='CPI/RS-adjusted Yearly Contribution Limit',xlab='Yearly Contribution Limit',ylab='Time')
plot(y=return$CPIC_limit,x=return$Time,main='CPIC-adjusted Yearly Contribution Limit',xlab='Yearly Contribution Limit',ylab='Time')
plot(y=return$PPI_limit,x=return$Time,main='PPI-adjusted Yearly Contribution Limit',xlab='Yearly Contribution Limit',ylab='Time')
```

## PCE_PR portfolio never reaches $1m from any starting time since 1959 Feb.


```{r, echo=FALSE, warning=FALSE, fig.height=3.5, fig.width=8}
# GR
return.temp = return %>% 
  gather(Portfolio, t, c('CPI_GR_t2','CPIC_GR_t2','PPI_GR_t2','RS_GR_t2')) %>%
  select(Portfolio, t)
p = ggplot(return.temp, aes(x=Portfolio, y=t), color=as.factor(Portfolio)) + 
    geom_boxplot(fill="white",color="steelblue")+
    xlab("Gross Return(GR) Portfolios")+
    ylab("Time to Reach $1m in Years")+
    theme_minimal()+
    labs(title="Boxplot of Time to Reach $1m for Different Gross Return(GR) Portfolios")+ 
    annotate("text", x=4, y=median(return.temp$t,na.rm = TRUE)+1, label= paste("Median:",round(median(return.temp$t,na.rm = TRUE),1),"years"),color="blue")+
    geom_hline(aes(yintercept = median(return.temp$t,na.rm = TRUE)),col='blue',lty=2,size=1)+
    stat_summary(fun.y=median, colour="red", geom="text", show_guide = FALSE, vjust=1.5, aes(label=round(..y.., digits=1)))+
    stat_summary(fun.y=median, colour="darkred", geom="point", size=3,show_guide = FALSE)+
    annotate("text", x=4, y=mean(return.temp$t,na.rm = TRUE)-1, label= paste("Mean:",round(mean(return.temp$t,na.rm = TRUE),1),"years"),color="red")+
    geom_hline(aes(yintercept = mean(return.temp$t,na.rm = TRUE)),col='red',lty=2,size=1)+
    scale_x_discrete(labels=c('CPI_GR','CPIC_GR','PPI_GR','RS_GR'))
print(p)

#PR
return.temp = return %>% 
  gather(Portfolio, t, c('CPI_PR_t2','CPIC_PR_t2','PPI_PR_t2')) %>%
  select(Portfolio, t)
p = ggplot(return.temp, aes(x=Portfolio, y=t), color=as.factor(Portfolio)) + 
    geom_boxplot(fill="white",color="steelblue")+
    xlab("Price Return(PR) Portfolios")+
    ylab("Time to Reach $1m in Years")+
    theme_minimal()+
    labs(title="Boxplot of Time to Reach $1m for Different Price Return(PR) Portfolios")+ 
    annotate("text", x=3.3, y=median(return.temp$t,na.rm = TRUE)-0.5, label= paste("Median:\n",round(median(return.temp$t,na.rm = TRUE),1),"years"),color="blue")+
    geom_hline(aes(yintercept = median(return.temp$t,na.rm = TRUE)),col='blue',lty=2,size=1)+
    stat_summary(fun.y=median, colour="red", geom="text", show_guide = FALSE, vjust=1.5, aes(label=round(..y.., digits=1)))+
    stat_summary(fun.y=median, colour="darkred", geom="point", size=3,show_guide = FALSE)+
    annotate("text", x=3.3, y=mean(return.temp$t,na.rm = TRUE)+0.5, label= paste("Mean:\n",round(mean(return.temp$t,na.rm = TRUE),1),"years"),color="red")+
    geom_hline(aes(yintercept = mean(return.temp$t,na.rm = TRUE)),col='red',lty=2,size=1)+
    scale_x_discrete(labels=c('CPI_PR','CPIC_PR','PPI_PR'))
print(p)
```

\newpage

```{r,echo=FALSE, fig.height=3.5, fig.width=8, echo=FALSE, warning=FALSE}
# GR
unfulfiled_t = data.frame("Portfolio"='a',
                          "Unfulfiled.start.time"=as.Date("2019-12-01"),
                          "Unfulfiled.end.time"=as.Date("2019-12-01"),stringsAsFactors=FALSE)
for (col in c(39,41,43,47)){
  unfulfiled_t=rbind(list(as.factor(colnames(return)[col]),
                          as.Date(return$Time[which(is.na(return[col]))[1]]),
                          as.Date("2019-12-01")), unfulfiled_t)
}
unfulfiled_t=unfulfiled_t[-nrow(unfulfiled_t),]
ggplot() + 
  geom_linerange(data=unfulfiled_t, aes(x=Portfolio, ymin=Unfulfiled.start.time, ymax=Unfulfiled.end.time),size=1.5)+
  coord_flip()+
  xlab("Portfolio")+
  ylab("Start time")+
  theme_minimal()+
  labs(title="Impossible Start Time for GR Portfolios to Reach $1m in 2020")+ 
  annotate("text", y=min(unfulfiled_t$Unfulfiled.start.time)+900, x=3.7,
           label=min(unfulfiled_t$Unfulfiled.start.time),color="blue")+
  annotate("point", y=min(unfulfiled_t$Unfulfiled.start.time), x=4, colour = "lightblue", size=5)+
  annotate("text", y=max(unfulfiled_t$Unfulfiled.start.time)+900, x=1.7,
           label=max(unfulfiled_t$Unfulfiled.start.time),color="blue")+
  annotate("point", y=max(unfulfiled_t$Unfulfiled.start.time), x=2, colour = "lightblue", size=5)

#PR
unfulfiled_t = data.frame("Portfolio"='a',
                          "Unfulfiled.start.time"=as.Date("2019-12-01"),
                          "Unfulfiled.end.time"=as.Date("2019-12-01"),stringsAsFactors=FALSE)
for (col in c(40,42,44,46)){
  unfulfiled_t=rbind(list(as.factor(colnames(return)[col]),
                          as.Date(return$Time[which(is.na(return[col]))[1]]),
                          as.Date("2019-12-01")), unfulfiled_t)
}
unfulfiled_t=unfulfiled_t[-nrow(unfulfiled_t),]
ggplot() + 
  geom_linerange(data=unfulfiled_t, aes(x=Portfolio, ymin=Unfulfiled.start.time, ymax=Unfulfiled.end.time),size=1.5)+
  coord_flip()+
  xlab("Portfolio")+
  ylab("Start time")+
  theme_minimal()+
  labs(title="Impossible Start Time for PR Portfolios to Reach $1m in 2020")+ 
  annotate("text", y=min(unfulfiled_t$Unfulfiled.start.time)+1200, x=3.7,
           label=min(unfulfiled_t$Unfulfiled.start.time),color="blue")+
  annotate("point", y=min(unfulfiled_t$Unfulfiled.start.time), x=4, colour = "lightblue", size=5)+
  annotate("text", y=max(unfulfiled_t$Unfulfiled.start.time)+1200, x=2.7, 
           label=max(unfulfiled_t$Unfulfiled.start.time),color="blue")+
  annotate("point", y=max(unfulfiled_t$Unfulfiled.start.time), x=3, colour = "lightblue", size=5)

```