---
title: "Simulation"
output: pdf_document
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
source("R/Simulation.R")
```

# Simple Bootstraping

## CPI-adjusted Gross Return

```{r, echo=FALSE}
Sim_Return_Plot(CPI_adjusted, "CPI_adjusted_GR", 1000, simfun = Spl_Boot)
```

## CPI-adjusted Price Return

```{r, echo=FALSE}
Sim_Return_Plot(CPI_adjusted, "CPI_adjusted_PR", 1000, simfun = Spl_Boot)
```

## CPIC-adjusted Gross Return

```{r, echo=FALSE}
Sim_Return_Plot(CPIC_adjusted, "CPIC_adjusted_GR", 1000, simfun = Spl_Boot)
```

## CPIC-adjusted Price Return

```{r, echo=FALSE}
Sim_Return_Plot(CPIC_adjusted, "CPIC_adjusted_PR", 1000, simfun = Spl_Boot)
```


## PPI-adjusted Gross Return

```{r, echo=FALSE}
Sim_Return_Plot(PPI_adjusted, "PPI_adjusted_GR", 1000, simfun = Spl_Boot)
```

## PPI-adjusted Price Return

```{r, echo=FALSE}
Sim_Return_Plot(PPI_adjusted, "PPI_adjusted_PR", 1000, simfun = Spl_Boot)
```

## PCE-adjusted Gross Return

```{r, echo=FALSE}
Sim_Return_Plot(PCE_adjusted, "PCE_adjusted_GR", 1000, simfun = Spl_Boot)
```

## PCE-adjusted Price Return

```{r, echo=FALSE}
Sim_Return_Plot(PCE_adjusted, "PCE_adjusted_PR", 1000, simfun = Spl_Boot)
```

## Robert Shiller-adjusted Return

```{r, echo=FALSE}
Sim_Return_Plot(RS_adjusted, "CPI_adjusted_Return", 1000, simfun = Spl_Boot)
```



# Block Bootstraping

## CPI-adjusted Gross Return

```{r, echo=FALSE}
Sim_Return_Plot(CPI_adjusted, "CPI_adjusted_GR", 1000, simfun = Blk_Boot)
```

## CPI-adjusted Price Return

```{r, echo=FALSE}
Sim_Return_Plot(CPI_adjusted, "CPI_adjusted_PR", 1000, simfun = Blk_Boot)
```

## CPIC-adjusted Gross Return

```{r, echo=FALSE}
Sim_Return_Plot(CPIC_adjusted, "CPIC_adjusted_GR", 1000, simfun = Blk_Boot)
```

## CPI-adjusted Price Return

```{r, echo=FALSE}
Sim_Return_Plot(CPIC_adjusted, "CPIC_adjusted_PR", 1000, simfun = Blk_Boot)
```

## PPI-adjusted Gross Return

```{r, echo=FALSE}
Sim_Return_Plot(PPI_adjusted, "PPI_adjusted_GR", 1000, simfun = Blk_Boot)
```

## CPI-adjusted Price Return

```{r, echo=FALSE}
Sim_Return_Plot(PPI_adjusted, "PPI_adjusted_PR", 1000, simfun = Blk_Boot)
```

## PCE-adjusted Gross Return

```{r, echo=FALSE}
Sim_Return_Plot(PCE_adjusted, "PCE_adjusted_GR", 1000, simfun = Blk_Boot)
```

## PCE-adjusted Price Return

```{r, echo=FALSE}
Sim_Return_Plot(PCE_adjusted, "PCE_adjusted_PR", 1000, simfun = Blk_Boot)
```

## Robert Shiller-adjusted Return

```{r, echo=FALSE}
Sim_Return_Plot(RS_adjusted, "CPI_adjusted_Return", 1000, simfun = Blk_Boot)
```


# Parametric Simulation -- LogNormal

## CPI-adjusted Gross Return

```{r, echo=FALSE}
Sim_Return_Plot(CPI_adjusted, "CPI_adjusted_GR", 1000, simfun = LN_Sim)
```

## CPI-adjusted Price Return

```{r, echo=FALSE}
Sim_Return_Plot(CPI_adjusted, "CPI_adjusted_PR", 1000, simfun = LN_Sim)
```

## CPIC-adjusted Gross Return

```{r, echo=FALSE}
Sim_Return_Plot(CPIC_adjusted, "CPIC_adjusted_GR", 1000, simfun = LN_Sim)
```

## CPIC-adjusted Price Return

```{r, echo=FALSE}
Sim_Return_Plot(CPIC_adjusted, "CPIC_adjusted_PR", 1000, simfun = LN_Sim)
```

## PPI-adjusted Gross Return

```{r, echo=FALSE}
Sim_Return_Plot(PPI_adjusted, "PPI_adjusted_GR", 1000, simfun = LN_Sim)
```

## PPI-adjusted Price Return

```{r, echo=FALSE}
Sim_Return_Plot(PPI_adjusted, "PPI_adjusted_PR", 1000, simfun = LN_Sim)
```

## PCE-adjusted Gross Return

```{r, echo=FALSE}
Sim_Return_Plot(PCE_adjusted, "PCE_adjusted_GR", 1000, simfun = LN_Sim)
```

## PCE-adjusted Price Return

```{r, echo=FALSE}
Sim_Return_Plot(PCE_adjusted, "PCE_adjusted_PR", 1000, simfun = LN_Sim)
```

## Robert Shiller-adjusted Return

```{r, echo=FALSE}
Sim_Return_Plot(RS_adjusted, "CPI_adjusted_Return", 1000, simfun = LN_Sim)
```


